Millions of Type Errors

### 2023-03-31 oopsla23/paper.tex
drop in telemetry text from google doc

>>>>

\subsection{What are we logging (from gdoc)}

\FILL{} change to long-form paragraphs, use more space for examples

\begin{description}
  \item[D1] metadata: uid, timestamp, session duration
  \item[D2] mode (strict / nonstrict / nocheck)
  \item[D3] a 1-bit reason for sending: did the user switch module?
    \subitem we get logs for two reasons: (1) switch module, (2) randomly with
    low probability after typechecking
  \item[D4] size of codebase: num files, num lines
  \item[D5] size of edit region
  \item[D6] overall type errors: for old and curr states, have: \# total, \# in module, and \# in region
  \item[D7] specific errors in the edit region: in the edit region, for old and
    curr states, how many type errors for each type error code? (There is 1
    edit region per log.)
    \subitem Examples, if the old state had one type-A error:
    \begin{itemize}
      \item If the latest edits touched it and fixed it, the next log says 1
        old type-A and 0 new type-A
      \item touched it, didn't fix => 1 old-A, 1 new-A
      \item didn't touch (fixed or not) => 0 old-A, 0 new-A
    \end{itemize}
  \item[D8] ForcedStrict "what if" errors for nocheck: specific errors for the
    edit region computed as if the user had strict mode enabled (even though they
    have no-check enabled)
    \subitem This info is available because it gives better hints for
    autocomplete, because it assumes precise types for ALL libraries (rather
    than Any)
  \item[D9] TooComplexErrors: project-wide count for this one specific error type
\end{description}

\subsection{what to study}

RQs + Analysis

\begin{itemize}
  \item How often does the typechecker give up with CodeTooComplex errors?
  \subitem [[ For context: the type system is designed to give useful feedback
    to all sorts of code, whether strict-typed, nonstrict-typed, or nocheck.
    But it has an escape hatch: CTC. Hope it's rarely needed. ]]
    \begin{itemize}
      \item Count the overall CTC errors (in each mode) for a coarse view of
        how rare they are. [D9 of course, D8 to compare nocheck code with
        strict-mode code via the same typechecker]
      \subitem Note: We can't look at the code to learn why the errors
        happened. Can only look for trends and tell the Roblox team where to
        look.
      %% \subitem Count CTC errors within each session to see if these errors are intermittent, or if they stick around. [D9 for the count, D8 for nocheck vs strict, D7 because CTC might show up in the edit region]
    \end{itemize}
  \item How helpful are highlighted regions?
    \begin{itemize}
      \item How often do edits in a highlighted region remove type errors? How
        often do they add new ones?
      \subitem [D7 compare specific old vs specific new]
      \subitem [may want to filter using D2 + D5]
      \item How often do edits outside the region remove type errors? How often do they add new ones?
      \subitem [D6 looking for old count vs. new count difference without a matching delta in the D7 counts]
      \item Which specific errors seem easy / hard to fix via highlights?
      \subitem [D6 removed by edit => easy; D6 survive edit => hard]
    \end{itemize}
  \item In what ways do types improve quality of life?
  \begin{itemize}
    \item Does nocheck code have many more latent errors than the others?
    \subitem [D8 latent errors come from the "what if" analysis]
  \end{itemize}
\end{itemize}

% What to study?
% 
% Sanity checks
% How large are edit regions? Plot the distribution.
% How long between logs? Long times & very short times are problematic.
% General Roblox trends
% How many sessions use strict / nonstrict / nocheck?
% Expect few strict/nonstrict, lots of nocheck
% How often does mode change within a session?
% Expect few of these b/c error reports are easy to ignore; non-intrusive
% If a down-switch happens, expect the current #errors to be high
% If an up-switch happens, hope to stay up
% How many CodeTooComplex errors arise, overall?
% Expect very few. Hopefully zero.
% How often do module switches happen?
% Expect not-too-many, with some logs in between each time
% Compare strict vs. nonstrict vs. nocheck
% Get median # errors across all snapshots.
% Expect nocheck median to be higher than the others.
% Plot #errors vs time, or vs codebase size
% Expect nocheck errors to keep going up. For other modes, #errors should stay low.
% General type error trends
% Plot the distribution of #errors across all edit regions
% Expect different counts based on error-kind. SEE BELOW
% Count specific errors that get fixed. For each kind of error, how many times is it old-in-region but not new-in-region? Compare to #old-in-region totals.
% Expect all errors to have a decent fix rate.
% If one error has a lower % than others, its error region may be misleading. Or, maybe it's hard to fix
% Type errors vs. edit region
% Compare #old vs. #new in region
% Expect a decrease, #old > #new
% If there's an increase, look for an explanation. Could be a huge edit region. Could be a 1-line region that we captured 2x in a row. Could be that this user is simply ignoring type errors.
% Compare [#old-in-region vs. (total #old - total #new)]
% If #old-in-region = 0, then hope the total # increases
% Decrease => nonlocal edits fixed an error => type error hilite may be inaccurate
% If #old-in-region > 0, then hope the total # decreases
% Increase (in region or otherwise) => made things worse
